Seed germination ecology of European mesic grasslands

### 2021-04-30 README.Rmd
intro 🌧️

# Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)

This repository stores all the information related to the manuscript *Seed germination ecology of species-rich mesic grasslands (Arrhenatheretalia)*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

>>>>

# Seed germination ecology of European mesic grasslands

This repository stores all the information related to the manuscript *Seed germination ecology of European mesic grasslands*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

### 2021-04-30 doc/manuscript.Rmd
intro 🌧️

---
title: "Seed germination ecology of species-rich mesic grasslands (*Arrhenatheretalia*)"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

@@

Eduardo Fernández-Pascual^1^ ^+^, Elías Afif Khouri^1^, Joaquín Ascaso^2^, Beatriz Morais^1^, Ramón Reiné^2^, Madalena Vaz^3^

^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^3^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal

@@

*Arrhenatherion*, plant regeneration, seed germination, seed morphology

@@

Iberian hay meadows [@RN2980] are seminatural grassland communities which developed in clearings of a formerly wooded landscape. They are species-rich mesophilous habitats, rarely fertilized, and occupy deep soils from clays to sands.  Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888]. Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation, and we can expect that seed morphology and germination timing will play a functional role in community assembly.

@@

To select species for inclusion in our analysis, we used a dataset of 118 vegetation plots from three Iberian regions with well-maintained mesic grasslands: 43 plots from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Spanish Pyrenees (Aragón). The dataset contains a relatively high diversity of mesic grassland species, as the plots were situated along a major stress gradient related to drought [@RN2980]: two Spanish regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893]. Furthermore, grasslands from the Pyrenees are closest to the central European optimum of mesic grassland vegetation (*Arrhenatheretalia elatioris* Tx. 1931) while Portuguese grasslands are in suboptimal areas at the limit of the distribution [@RN4893].

Each plot was square (25-100 m^2 area) and was situated in the central part of a meadow avoiding the margins. The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year). Vegetation in the plots was sampled in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots were recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names in this manuscript follow the nomenclature of the Euro+Med plant database (https://www.emplantbase.org/home.html). Using all the plots, we calculated the cumulative abundance of each species in the entire area. We removed 208 species with cumulative abundance values below 2%, considering them transient species that might have been recorded by chance and do not represent the core mesic grasslands flora. We used the remaining 117 species as the core list to retrieve seed germination data for this article.

## Species collection and morphology

From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016-2018. To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678]. We distributed the seeds on the scanner transparent glass, in a 10 x 10 grid. For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds. In the case of the black background, we covered samples with a black box to avoid interference from environmental light. For the white background, we used the scanner cover. We digitized the obtained images and stored them in JPEG format (Joint Photographic Experts Group). We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708]. The program calculates several biometric parameters for each seed on the sample, among these, seed length and width. Using these values, we calculated seed shape as seed length divided by seed width. Additionally, we obtained seed mass for the study species from Kew's Seed Information Database [@RN2987].

@@

To describe seed germination requirements, we germinated seeds of the different species following a protocol that accounts for the effect of germination temperature as a main physiological driver of germination phenology [@RN2249]. We used three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of the Fabaceae, the treatment consisted in scarification by chipping the seed coat with a scalpel, to remove physical dormancy [@RN3214]. For the rest of the families, we used an incubation in gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove any potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12 hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy. We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.

In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. The combined dataset, including our own experimental data and the records from ENSCOBASE, contained `r MSrecords` records (i.e. germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions). There were `r MSspecies` species in the dataset. Overall, `r MSseeds` seeds had been used in the experiments. The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments went from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records and kept in total darkness in `r MSlightN` records. Experiments had been performed with unstratified seeds (i.e. not subjected to a previous dormancy-breaking incubation) in `r MSstratN` records and with stratified seeds (i.e. subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records. Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.

@@

The dispersal unit in the dominant grass family (Poaceae) is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis (a dry, indehiscent and monospermic fruit) and several surrounding bracts (i.e. modified leaves): the lemma, the palea and, in many species, two glumes. This structure is often accompanied by awns or hairs that aid in dispersal (**Fig. 1b**). The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm). In the subordinate family of the legumes (Fabaceae), the dispersal unit of most species is the seed itself (**Fig. 1c**). Fabaceae generally are hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy). In some Fabaceae, dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs. `r MSlegWidth` mm wide). In many of the most frequent minor families, such as the Asteraceae, the Ranunculaceae, the Lamiaceae and the Dipsacaceae, the dispersal unit is the achene, i.e. a dry, indehiscent and monospermic fruit. The achenes of the grassland Asteraceae tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig. 1f**). In the Ranunculaceae, the achenes are usually flattened and sometimes beaked (**Fig. 1g**). The achenes of the Lamiaceae are hardened nutlets (**Fig. 1h**). In the genus *Knautia* (Dipsacaceae), the achene is covered by hairs (**Fig. 1i**). In the Cyperaceae, the dispersal unit is an achene surrounded by a perigynium which aids in dispersal by water (**Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). The dispersal unit of the Apiaceae is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**) or the Plantaginaceae (**Fig. 1n**). The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig. 1ñ**). When comparing the morphology of all species (**Fig. 2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.

@@

When considering the full set of mesic grassland species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig. 3**). The germination of grassland seeds was increased by scarification, stratification, GA~3~, alternating temperatures, and light. Temperature had a negative effect, indicating a trend towards higher germination at lower temperatures. Averaging the whole dataset, the highest germination proportions were achieved at 20ºC. Between 0ºC and 20ºC, germination proportions increased steadily with increasing temperatures. Above 20ºC, germination declined more sharply, and grassland seeds rarely germinated at 30ºC.

Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation. In the dominant group of the Poaceae (**Fig. 3**), no effect was found for stratification, GA~3~ or temperature. Stratification had a negative effect on germination. The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect. In the subordinate group of the Fabaceae (**Fig. 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light. Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20ºC than Poaceae species did. Fabaceae seed germination also responded negatively to alternating temperatures. Finally, for the group of the minor families (**Fig. 3**), the main cues having a positive effect on germination were stratification, GA~3~. alternating temperatures and light. These species did not respond to scarification nor to temperature.

To describe the effects of the random factors, we will refer only to the full model that included all species in the dataset. The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`), the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`), and species identity (mean = `r MSspecies`, CI = `r MSspeciesl` - `r MSspeciesh`). The phylogenetic signal in the germination responses was relatively high (lambda = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).

@@

@@

E.F.P. conceived the study, performed the analyses and wrote the manuscript. All authors contributed data, revised the manuscript, and approved the final version.

@@

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic grasslands: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* subsp. *despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (ñ) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
knitr::include_graphics("../results/figures/seeds.png")
```

@@

```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on seed germination of mesic grassland species. Dots indicate the posterior mean of the effect size, and whiskers its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands"}
knitr::include_graphics("../results/figures/mcmc.png")
```

>>>>

---
title: "Seed germination ecology of European mesic grasslands"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

@@

Eduardo Fernández-Pascual^1^ ^+^, Madalena Vaz^2^, Beatriz Morais^1^, Ramón Reiné^3^, Joaquín Ascaso^3^, Elías Afif Khouri^1^, Angelino Carta^4^

^1^ Departamento de Biología de Organismos y Sistemas, Universidad de Oviedo, Oviedo/Uviéu, Spain; ^2^ Banco Português de Germoplasma Vegetal, Instituto Nacional de Investigação Agrária e Veterinária (INIAV), Braga, Portugal; ^3^ Departamento de Ciencias Agrarias y Medio Natural, Universidad de Zaragoza, Huesca, Spain; ^4^ Dipartimento di Biologia, Botany Unit, University of Pisa, Pisa, Italy

@@

*Arrhenatheretalia*, plant regeneration, seed germination, seed morphology, species-rich meadows, hay meadows, mesic pastures

@@

European mesic grasslands are semi-natural open habitats that occupy moderately fertile and well-drained soils [@RN4698]. These plant communities developed by anthropogenic intervention in clearings of a formerly wooded landscape and are maintained by agricultural practices, contributing to the diversity and conservation value of European landscapes [@RN4996]. Traditional phytosociological classification of mesic grasslands put emphasis on the separation between mesic meadows used for hay making versus mesic pastures maintained by grazing, but a recent revision at the European level showed that the main driver of variation in species composition is the intensity rather than the type of management [@RN4893]. Indeed, changes in the intensity of management, including land abandonment and agricultural intensification, are threatening the maintenance of mesic grasslands in large parts of Europe [@RN4997]. For this reason the European Habitats Directive (92/43/EEC) has included in its lists of habitats of conservation interest those species-rich meadows that are traditionally managed by one or two annual cuts and light grazing.

One of the reasons for the high conservation interest of traditional mesic grasslands is their high species richness, and for this same reason they have been highlighted as a valuable source of natural seed materials to be used in ecological restoration, rewilding and urban greening [@RN4894]. However, a lack of knowledge about species germination traits has already been identified as a bottleneck [@RN2241] hampering the development of a competitive native seed industry [@RN4665] in Europe. Understanding seed germination is part of the practical scientific framework needed to tackle large-scale ecological restoration challenges [@RN3507] and to maintain ex situ collections of plant genetic resources [@RN3763]. When such a knowledge framework is available, ambitious regional schemes of seed-based landscape restoration can be designed [@RN4666].

In the case of mesic grasslands, a concerted investigation of their seed germination ecology is missing, despite their conservation and restoration importance. Most of the available information is scattered in single-species studies [@RN2236] or using a limited range of germination conditions [@RN4894]. In this article, we present for the first time a synthesis of germination ecology in European mesic grasslands. To do so, we combined a newly-generated dataset with germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. To the resulting dataset containing `r MSrecords` germination records from `r MScountries` European countries we applied Bayesian meta-analysis [@RN5000] to test the seed germination response to environmental cues including scarification, stratification, temperature, alternating temperature and light. In addition, we also analyzed seed morphology (mass and shape) and the differences between the main functional groups of mesic grasslands: grasses (Poaceae), legumes (Fabaceae) and the other families.

@@

To create a species list of representative mesic grassland species for inclusion in our analysis, we used a dataset of 118 vegetation relevés from three Western European regions with well-maintained mesic grasslands: 43 from Northern Portugal (Trás-os-Montes), 25 from the Cantabrian Mountains of Spain (Parque Nacional Picos de Europa) and 50 from the Pyrenees. This dataset represents a sample of mesic grassland diversity, as the plots were situated along a major stress gradient related to drought [@RN2980]: two regions with a Temperate macroclimate, and Northern Portugal which is transitional between the Temperate and Mediterranean macroclimates [@RN4893]. Furthermore, grasslands from the Pyrenees are closest to the central European optimum of mesic grassland vegetation (*Arrhenatheretalia elatioris* Tx. 1931) while Portuguese grasslands are in suboptimal areas at the limit of the distribution [@RN4893]. Finally, traditional management of mesic grasslands is relatively well preserved in this area within the European context [@RN4998; @RN4999].

Each relevé had been conducted in a square area (25-100 m^2^ area) situated in the central part of a meadow avoiding the margins. The chosen meadows are maintained by traditional agricultural practices (mowing for hay-making once or twice per year and light grazing). Vegetation in the plots had been sampled in 2016-2017, at the peak of plant development, just before mowing. All vascular plant species in the plots had been recorded and assigned an abundance value using the transformation of the Braun-Blanquet scale to coverage (+ = 0.1%, 1 = 5%, 2 = 17.5%, 3 = 37.5%, 4 = 62.5% and 5 = 87.5%). All plant names in this manuscript follow the nomenclature of the Euro+Med plant database (https://www.emplantbase.org/home.html). Using all the relevés, we calculated the cumulative abundance of each species in the entire area. We removed 208 species with cumulative abundance values below 2%, considering them transient species that might have been recorded by chance and do not represent the core mesic grasslands flora. We used the remaining 117 species as the core list to retrieve seed germination data for this article.

## Seed collection and morphology

From the vegetation plots described above, we collected dispersal units (hereafter called seeds) during the dispersal seasons of 2016, 2017 and 2018. To describe seed morphology, we acquired images of 100-seed samples of each species using a flatbed scanner (Brother LC985) with a resolution of 200 dpi and a scanning area of 1024 x 1024 pixels [@RN2678]. We distributed the seeds on the scanner transparent glass, in a 10 x 10 grid. For each sample, and without moving the seeds, we repeated the scans with black and white backgrounds. In the case of the black background, we covered samples with a black box to avoid interference from environmental light. For the white background, we used the scanner cover. We digitized the obtained images and stored them in JPEG format (Joint Photographic Experts Group). We processed the scanned images using *ImageJ*, an open-source image processing program designed for scientific multidimensional images [@RN4708]. The program calculates several biometric parameters for each seed on the sample, among these, seed length and width. Using these values, we calculated seed shape as seed length divided by seed width. Additionally, we obtained seed mass for the study species from Kew's Seed Information Database [@RN2987].

@@

We germinated seeds of the different species using three germination treatments to determine the effect of temperatures representative for the study area: 14/4 ºC representing the capacity of freshly-dispersed seeds to germinate at cool temperatures of spring and autumn, 22/12 ºC as the optimal summer temperature, and 30/20 ºC as sun-heated soil, e.g. soil exposed to sun after hay cutting. Additionally, we compared, at each of these temperature regimes, the germination of fresh seeds versus seeds subjected to a dormancy-breaking treatment. In the case of the Fabaceae, the treatment consisted in scarification by chipping the seed coat with a scalpel, to remove physical dormancy [@RN3214]. For the rest of the families, we used gibberellic acid GA~3~ (0,0645 mM) in darkness during 24h, as a treatment to remove potential physiological seed dormancy [@RN4976]. For each species and treatment, we sowed four Petri dishes with 25 seeds each. The germination substrate was 1% distilled water - agar. We sealed dishes with Parafilm to prevent desiccation. Trials took place in a germination chamber (KBW 400, Binder GmbH, Tuttlingen, Germany) with a 12/12 hour photoperiod (the light period corresponding to the higher temperature). Experiments lasted for four weeks, with germination scoring once per week. The germination criterion was 2 mm radicle emergence. After four weeks, we cut the seeds that failed to germinate and examined them under a magnifying glass. We classified them as normal when the embryo was visible and firm, empty when they lacked an embryo, and contaminated when they were moldy. We only considered normal seeds to calculate germination proportions and conduct subsequent analyses.

In addition to this experimental data, we retrieved seed germination records from ENSCOBASE, the seed germination database of the European Native Seed Conservation Network (http://enscobase.maich.gr/index.tml); and the *SylvanSeeds* database of seed germination records for nemoral biome species [@RN4968]. The combined dataset, including our own experimental data and the records from ENSCOBASE and *SylvanSeeds*, contained `r MSrecords` germination records (i.e. germination proportions for a given seed lot of a species, recorded in a set of laboratory experimental conditions) from `r MScountries` European countries. There were `r MSspeciesN` species in the dataset. Overall, `r MSseedsN` seeds had been used in the experiments. The range of experimental germination temperatures (weighted average of the daily thermoperiod) which had been used in the experiments spans from `r MSminT` to `r MSmaxT` ºC, with `r MSaltN` records of constant temperatures (i.e. experiments that used the same temperature during all their duration) and `r MSaltY` of alternating temperatures (i.e. experiments where different temperatures were applied during the day and the night, in diurnal cycles). Seeds had been exposed to light during some part of the diurnal cycle in `r MSlightY` records or kept in total darkness in `r MSlightN` records. Experiments had been performed with unstratified seeds (i.e. not subjected to a previous dormancy-breaking incubation) in `r MSstratN` records and with stratified seeds (i.e. subjected to previous incubation in dormancy-breaking conditions, including treatments of wet incubation under cold, warm and combinations of cold and warm conditions) in `r MSstratY` records. Finally, there were `r MSgaY` records where GA~3~ had been applied, and `r MSscarY` records where seeds had been scarified.

@@

The dispersal unit in the dominant grass family (Poaceae) is generally the floret (**Fig. 1a**), a composite structure made up of the caryopsis (a dry, indehiscent and monospermic fruit) and several surrounding bracts (i.e. modified leaves): the lemma, the palea and, in many species, two glumes. This structure is often accompanied by awns or hairs that aid in dispersal (**Fig. 1b**). The mass of these dispersal units in grassland grasses ranges from `r MSpoaMin` to `r MSpoaMax` mg, and they tend to be longer (`r MSpoaLength` mm on average) than they are wide (`r MSpoaWidth` mm). In the subordinate family of the legumes (Fabaceae), the dispersal unit of most species is the seed itself (**Fig. 1c**). Fabaceae generally are hard-seeded, i.e. they have a water-impermeable seed coat which needs to become permeable before germination can happen (i.e. physical seed dormancy). In some Fabaceae, dispersal units are more complex, including indehiscent monospermic fruits (e.g. *Onobrychis*, **Fig. 1d**) or indehiscent monospermic fruit fragments, i.e. loments (e.g. *Ornithopus*, **Fig. 1e**). In mesic grasslands, the dispersal units of the Fabaceae are generally heavier than those of the Poaceae, ranging from `r MSlegMin` to `r MSlegMax` mg; and they are also rounder (`r MSlegLength` mm long vs. `r MSlegWidth` mm wide). In many of the most frequent minor families, such as the Asteraceae, the Ranunculaceae, the Lamiaceae and the Dipsacaceae, the dispersal unit is the achene, i.e. a dry, indehiscent and monospermic fruit. In the case of the Asteraceae, these achenes tend to be cypselas, wearing a pappus that aids in dispersal by wind (**Fig. 1f**). In the Ranunculaceae, the achenes are usually flattened and sometimes beaked (**Fig. 1g**). The achenes of the Lamiaceae are hardened nutlets (**Fig. 1h**). In the genus *Knautia* (Dipsacaceae), the achene is covered by hairs (**Fig. 1i**). In the Cyperaceae, the dispersal unit is an achene surrounded by a perigynium which aids in dispersal by water (**Fig. 1j**). In the genus *Sanguisorba* (Rosaceae), the dispersal unit is the urn-shaped receptacle containing one to three achenes (**Fig. 1k**). The dispersal unit of the Apiaceae is the mericarp (**Fig. 1l**), an indehiscent monospermic fragment of the fruit. In some other minor families the dispersal unit is the seed itself, such as in the Caryophyllaceae (**Fig. 1m**) or the Plantaginaceae (**Fig. 1n**). The hemiparasitic species of *Rhinanthus* (Orobanchaceae) disperse by winged seeds (**Fig. 1o**). When comparing the morphology of all species (**Fig. 2**), the Poaceae appear as having relatively lighter and more elongated dispersal units, in contrast with the Fabaceae, which tend to have rounder and heavier seeds.

@@

When considering the full set of mesic grassland species, all of the six studied germination cues had a significant effect on final germination proportions (**Fig. 3**). The germination of grassland seeds was positively associated with scarification, stratification, GA~3~, alternating temperatures, and light. Average temperature had a negative effect, indicating a trend towards higher germination at lower temperatures. Averaging the whole dataset, the highest germination proportions were achieved at 20 ºC. Between 0 ºC and 20 ºC, germination proportions increased steadily with increasing temperatures. Above 20 ºC, germination declined more sharply, and grassland seeds rarely germinated at 30 ºC.

Some differences became apparent when dividing the dataset in the major floristic groups that compose mesic grassland vegetation (i.e. (i) the Poaceae as dominant and structural species, (ii) the Fabaceae as the second largest group, and (iii) the rest of the families). In the dominant group of the Poaceae (**Fig. 3**), no effect was found for stratification, GA~3~ or temperature. Stratification had a negative effect on germination. The major drivers of Poaceae germination appeared to be alternating temperatures and light, with both having a positive effect. In the group of the Fabaceae (**Fig. 3**), the largest positive effect on germination was produced by scarification, with no effect of stratification, GA~3~ or light. Temperature had a negative effect, and in fact Fabaceae species had higher germination proportion at temperatures under 20 ºC. Fabaceae seed germination also responded negatively to alternating temperatures. Finally, the remaining families (**Fig. 3**), the main cues having a positive effect on germination were stratification, GA~3~. alternating temperatures and light. These species did not respond to scarification nor to average temperature.

To describe the effects of the random factors, we will refer only to the full model that included all species in the dataset. The strongest effect was that of the phylogeny (mean = `r MSphylo`, CI = `r MSphylol` - `r MSphyloh`), followed by the source of the data (mean = `r MSsource`, CI = `r MSsourcel` - `r MSsourceh`) and the seed lot (mean = `r MSpopulation`, CI = `r MSpopulationl` - `r MSpopulationh`). The phylogenetic signal in the germination responses was relatively high (lambda = `r MSlambda`, CI = `r MSlambdal` - `r MSlambdah`).

@@

* Mowing for haymaking thus is a predictable perturbation that occurs approximately at the same time every year, and which is expected to contribute to the high diversity of this habitats [@RN4887; @RN4888]. Hay meadow plants must establish, grow and reproduce in the time frame determined by this perturbation.

* Question of clonal regeneration.

* Discuss second axis = geophytes vs hemicriptófitos?

@@

All authors contributed data. E.F.P. conceived the study and performed the analyses. E.F.P. wrote the manuscript with help from A.C. All authors revised the manuscript, and approved the final version.

@@

```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Diversity of dispersal units in mesic grasslands: (a) Floret of *Cynosurus cristatus* (Poaceae); (b) floret of *Arrhenatherum elatius* (Poaceae); (c) seed of *Lathyrus pratensis* (Fabaceae); (d) legume of *Onobrychis viciifolia* (Fabaceae); (e) loment fragment of *Ornithopus perpusillus* (Fabaceae); (f) achene with pappus of *Centaurea scabiosa* (Asteraceae); (g) achene of *Ranunculus acris* subsp. *despectus* (Ranunculaceae); (h) nutlet of *Prunella grandiflora* (Lamiaceae); (i) achene of *Knautia nevadencis* (Dipsacaceae); (j) perigynium of *Carex binervis* (Cyperaceae); (k) receptacle of *Sanguisorba minor* (Rosaceae); (l) mericarp of *Carum verticillatum* (Apiaceae); (m) seed of *Cerastium fontanum* (Caryophyllaceae); (n) seed of *Plantago lanceolata* (Plantaginaceae); (o) seed of *Rhinanthus angustifolius* (Orobanchaceae)."}
knitr::include_graphics("../results/figures/seeds.png")
```

@@

```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Effect of the germination environment on seed germination of mesic grassland species. Dots indicate the posterior mean of the effect size, and whiskers its 95 % credible interval. The line of zero effect is shown. When the credible intervals overlap with the zero-effect line, the effect is not significant. The figure shows the results of a model including all species, and of models including only the three main floristic groups of mesic grasslands (i.e. the Poacaea, the Fabaceae and the remaining families)."}
knitr::include_graphics("../results/figures/mcmc.png")
```
